---
title: "Grammar of Graphics"
author:
- "Tilina Alzaben"
output:
  html_document: default
  pdf_document: default
---

# Resisting the First Order for Categories

For this problem, we'll work with the `starwars` dataset that is included in the [`dplyr` package](https://dplyr.tidyverse.org/reference/starwars.html).

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
data("starwars")
starwars_character_films <- starwars %>%
  dplyr::select(name, films) %>%
  unnest(films)
```

We will focus on the `films` variable, which indicates the film a particular character appeared in. Specifically, we will demonstrate how to reorder categories on graphs in `R` which particularly useful when visualizing categorical variables. First, make a bar plot of the `films` variable. 

```{r}
starwars_character_films %>%
  ggplot(aes(x = fct_rev(fct_infreq(films)))) +
  geom_bar(fill = "lightblue") +
  labs(
    title = "Distribution of Characters in Star Wars",
    subtitle = "(ordered most to least frequent)",
    y = "Count", x = "Film"
  ) +
  coord_flip()
```

Recreate the plot, but this time, order the categories in a logical order (such as episode order: The Phantom Menace, Attack of the Clones, ..., The Force Awakens; or release date order: A New Hope, The Empire Strikes Back, ..., The Force Awakens).

```{r}
starwars_character_films %>%
  ggplot(aes(x = fct_rev(fct_relevel(
    films,
    "The Phantom Menace",
    "Attack of the Clones",
    "Revenge of the Sith",
    "A New Hope",
    "The Empire Strikes Back",
    "Return of the Jedi",
    "The Force Awakens"
  )))) +
  geom_bar(fill = "lightblue") +
  labs(
    title = "Distribution of Characters in Star Wars",
    subtitle = "(in chronological order)",
    y = "Count", x = "Film"
  ) +
  coord_flip()
```

Recreate the plot from the previous part, but this time, rename the categories so that they use the following film abbreviations:  I, II, III, IV, V, VI, VII.

```{r}
starwars_character_films <- starwars_character_films %>%
  mutate(films = fct_relevel(
    films,
    "The Phantom Menace",
    "Attack of the Clones",
    "Revenge of the Sith",
    "A New Hope",
    "The Empire Strikes Back",
    "Return of the Jedi",
    "The Force Awakens"
  ))


starwars_character_films %>%
  ggplot(aes(x = fct_rev(fct_recode(films,
    "I" = "The Phantom Menace",
    "II" = "Attack of the Clones",
    "III" = "Revenge of the Sith",
    "IV" = "A New Hope",
    "V" = "The Empire Strikes Back",
    "VI" = "Return of the Jedi",
    "VII" = "The Force Awakens"
  )))) +
  geom_bar(fill = "lightblue") +
  labs(
    title = "Distribution of Characters in Star Wars",
    subtitle = "(in chronological order)",
    y = "Count", x = "Film"
  ) +
  coord_flip()
```


# UFO Sightings in Pittsburgh

For this problem with a dataset via the #TidyTuesday project containing [information about UFO sightings](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-06-25). 

```{r}
library(tidyverse)
ufo_sightings <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv")
```


Create a __single graph__ that displays the __marginal distribution__ of `country`.

```{r, message = FALSE, warning = FALSE}
ufo_sightings %>%
  ggplot(aes(x = fct_infreq(country))) +
  geom_bar(fill = "lightgreen") +
  labs(
    title = "Distribution of UFO Sighting Countires",
    x = "Country", y = "UFO Sightings"
  )
```


Create a summarized dataset that contains the following information about `ufo_shape` for UFO sightings in Pittsburgh (note: use lower-case `pittsburgh`):

- `ufo_shape`: the UFO shape (this dataset should have one row for each shape)
- `prop`: proportion of Pittsburgh UFO sightings with the particular UFO shape
- `se`: standard error for UFO shape proportion estimate
- `lower`: lower bound for 95% CI
- `upper`: upper bound for 95% CI

Make sure to only look at UFO sightings in Pittsburgh and remove any sightings where the `ufo_shape` is missing.

```{r}
ufo_sightings_summary <- ufo_sightings %>%
  filter(str_to_lower(city_area) == "pittsburgh" & !is.na(ufo_shape)) %>%
  group_by(ufo_shape) %>%
  summarize(
    count = n(),
    .group = "drop"
  ) %>%
  mutate(
    total = sum(count),
    prop = count / total,
    se = sqrt(prop * (1 - prop) / total),
    lower = prop - 2 * se,
    upper = prop + 2 * se
  ) %>%
  select(ufo_shape, prop, se, lower, upper)

ufo_sightings_summary
```

Create a __single graph__ that displays the __marginal distribution__ of `ufo_shape` for UFO sightings in Pittsburgh, using the dataset you constructed above. Additionally, make sure your graph has the following:

- add 95% CI to your display,
- order the `ufo_shape` category levels by frequency in your display,
- make sure the `ufo_shape` labels are easy to read and that your graph is labeled appropriately.

```{r}
ufo_sightings_summary %>%
  ggplot(aes(x = fct_reorder(ufo_shape, prop))) +
  geom_bar(aes(y = prop), stat = "identity", fill = "lightgreen") +
  coord_flip() +
  geom_errorbar(
    aes(
      ymin = lower,
      ymax = upper
    ),
    color = "darkgreen"
  ) +
  labs(
    title = "Frequency of Different UFO Shapes Seen in Pittsburgh",
    subtitle = "(ordered by most to least frequent)",
    x = "UFO Shape", y = "Sightings"
  )
```

In the previous part, we ignored the issue of multiple testing. You can find that there are 120 different pairwise comparisons ($120 = \binom{16}{2} =$ `ncol(combn(16,2))`) you could make from the graph above. As a result, if we attempt to make multiple pairwise comparisons from a single graph with 95% CIs, our chance of making a Type 1 error is greater than 5%. 

In general, $\alpha$-level CIs are constructed using the Normal quantile $z_{1 - \alpha / 2}$. When we construct 95% CIs, $\alpha = 0.05$, and thus we use $z_{1 - \alpha / 2} = z_{0.975}$ (computed with `qnorm(0.975)`).

Recreate the figure but update your CIs with new CIs following the [Bonferroni correction](https://en.wikipedia.org/wiki/Bonferroni_correction).

```{r}
ufo_sightings_summary %>%
  mutate(lower = pmax(0, prop - qnorm(1 - (0.05 / 120) / 2) * se), 
         upper = pmin(1, prop + qnorm(1 - (0.05 / 120) / 2) * se)) %>%
  ggplot(aes(x = fct_reorder(ufo_shape, prop))) +
  geom_bar(aes(y = prop), stat = "identity", fill = "lightgreen") +
  coord_flip() +
  geom_errorbar(
    aes(
      ymin = lower,
      ymax = upper
    ),
    color = "darkgreen"
  ) +
  labs(
    title = "Frequency of Different UFO Shapes Seen in Pittsburgh",
    subtitle = "(ordered by most to least frequent)",
    x = "UFO Shape", y = "Sightings"
  )
```
